### 2016-09-20 Thesis/DCG_sb-wi-thesis-template/intro/UPS.tex
added proteasome chapter
proteasome additions final edits after erin

Adapted in part from: R.S. Marshall*, D.C. Gemperline*, and R.D. Vierstra (2015). Purification of 26S proteasomes and their sub-complexes from plants. \textit{Methods Mol. Biol.} DOI: 10.1007/978-1-4939-6533-5

\noindent
*These authors contributed equally to the \textit{Methods Mol. Biol.} work.

\noindent
This chapter has expanded sections on the proteasomes role in plant biology, proteasome assembly, and proteasome isotypes.

\section{The Ubiquitin 26S Proteasome System (UPS)}
	Selective proteolysis in plants plays a critical role in both regulating growth and development, and maintaining cellular homeostasis \citep{nelson14, smalle04, vierstra93, vierstra09}.  One of the principle pathways for protein degradation in plants and other eukaryotes is the ubiquitin-26S proteasome system (UPS), which involves the attachment of polyubiquitin chains to target proteins followed by their recognition and degradation by the 26S proteasome, an exquisitely designed proteolytic machine \citep{bhattacharyya14, finley09, vierstra09}.  The UPS is highly conserved across all eukaryotes; it was first elucidated by elegant work in rabbit reticulocyte lysates \citep{ciechanover80, ciechanover80-frAQB, etlinger77, hershko80, wilkinson80}, and was subsequently identified in other animals, yeast and plants \citep{ciechanover84, finley84, finley87, glotzer91, hochstrasser91, shanklin87}.  Ubiquitin conjugation to target proteins is accomplished through a highly polymorphic, ATP-dependent cascade involving the sequential action of three enzyme classes, termed the E1 ubiquitin-activating enzymes, E2 ubiquitin-conjugating enzymes, and E3 ubiquitin-protein ligases \citep{berndsen14, smalle04, vierstra09} (\ref{fig:upscycle}).  Selectivity in ubiquitylation is typically driven by the E3 family, which has dramatically expanded during plant evolution to include well over a thousand variants in Arabidopsis and other plant species \citep{hua13, hua11}. Indeed 6\% of the Arabidopsis proteome is devoted to UPS \citep{smalle04}.

@@

	Additionally, the UPS also aides in the regulation of key plant hormonal responses including auxin, abscisic acid, jasmonic acid (JA), giberellins (GA), and ethylene. (reviewed in \citep{santner10}). Key hormone response factors are degraded by a variety of E3 complexes including TIR1 involved in auxin regulation \citep{dharmasiri05}, COI1, involved JA sensing \citep{katsir08}, and GID1 involved in GA responses \citep{murase08}. An elegant mechanism requires the direct binding of these key hormones either to their respective E3 complexes, or key proteins associated with these E3’s to promote the efficient ubiquitylation and subsequent degradation of their hormone response factors \citep{shabek14}.

@@

\section{The 26S Proteasome}
	The 26S proteasome is the central enzymatic effector of the UPS, and is a 2.5 MDa particle located in the cytosol and nucleus of eukaryotic cells.  It is composed of two functionally distinct sub-complexes; the 20S core protease (CP) that houses the proteolytic active sites, and the 19S regulatory particle (RP) that recognizes appropriate substrates (\ref{fig:proteasome} A and B; \citep{bhattacharyya14, finley09, lander12, lasker12, unverdorben14}).

@@

The CP has a barrel shape generated by four stacked hetero-heptameric rings, which contain seven $\alpha$-subunits or seven $\beta$-subunits (termed PAA-PAG and PBA-PBG, respectively, in \textit{Arabidopsis}) in an $\alpha$1-7/$\beta$1-7/$\beta$1-7/$\alpha$1-7 configuration (\ref{fig:cpdetails} A (side view- stacked rings), C (top view- $\alpha$-subunits) and D (top view- $\beta$-subunits).

@@

Upon assembly, a central chamber is formed at the $\beta$-ring interface that houses six peptidase catalytic sites provided by the $\beta$1 (PBA), $\beta$2 (PBB), and $\beta$5 (PBE) subunits \citep{arendt97, heinemeyer97} (\ref{fig:cpdetails} B and E).  The active sites involve a catalytic triad, one residue of which is an N-terminal threonine that becomes exposed during CP assembly.  Collectively these peptidases can cleave a broad range of protein sequences with peptidylglutamyl-peptide hydrolyzing (PGPH) ($\beta$1), trypsin-like ($\beta$2), and chymotryptic-like ($\beta$5) activities \citep{arendt97, groll99} (\ref{fig:cpdetails} E).  The $\alpha$-rings create two antechambers with narrow opposing axial pores that are gated by extensions at the N-terminus of several subunits (\ref{fig:cpdetails} B) \citep{groll00, ruschak10}.  Through this distinctive architecture, the CP acts as a self-compartmentalized protease that will only degrade polypeptides that are deliberately recognized, unfolded, and imported into the $\beta$-ring chamber.

	The CP is capped at one or both ends by the RP, which sits on top of the axial pores.  The RP provides activities for recognition of ubiquitylated proteins, substrate unfolding and import, and release of the ubiquitin moieties before substrate degradation (\ref{fig:proteasome}).  Its binding to the CP is stabilized by ATP, which is thus a necessary ingredient for purifying intact 26S proteasomes \citep{smith05}. The RP itself consists of two sub-complexes; the base, which contains a hexameric ring of AAA-ATPases (RPT1-6) plus two non-ATPase subunits, RPN1 and RPN2; and the lid, which is composed of an additional 11 non-ATPase subunits, RPN3, RPN5-13 and DSS1/SEM1 (\ref{fig:proteasome}; \citep{bhattacharyya14, book10, finley09, glickman98-c8Wsa, russell13}.  This lid/base demarcation was first revealed by the absence of lid subunits in proteasomes isolated from a \textit{Δrpn10} yeast deletion strain, and it was hence thought that RPN10 helps enforce binding of the lid to the base \citep{glickman98}.  However, more recent structural studies have demonstrated that RPN10 has a more indirect stabilizing role via its interaction with RPN9 \citep{lander12}.  The ring of RPT subunits in the base promotes substrate unfolding through ATP hydrolysis, and gates the $\alpha$-ring axial pores through repositioning of the CP $\alpha$-subunit extensions \citep{köhler01, rabl08, smith05}.  The N-terminal regions from proximal RPT pairs intertwine to create three spokes onto which most RPN subunits are scaffolded (\ref{fig:rpdetails} A; \citep{beck12}).

@@

Several RPT subunits harbor C-terminal \textit{Hydrophobic-Tyrosine-Any} amino acid (HbYX) motifs important for their binding to the $\alpha$-subunits in the CP. These C-terminal HbYX and HbYX-like extensions typically intercalate between adjacent $\alpha$-subunits (\ref{fig:rpdetails} B). The RPN6 subunit acts as a molecular clamp to tether the RP onto the CP (\ref{fig:rpdetails} A) \citep{pathare12}.

\section{Proteasome Purification Strategies}
	Even before the realization that the 26S proteasome is a protease, sub-particles of the complexes were described.  The first reports of proteasomes used avian erythroblast preparations enriched by differential ultracentrifugation followed by fractionation through a sucrose gradient \citep{schmid84}.  These 20S fractions isolated in the absence of added ATP were found to inhibit mRNA translation in a cell-free system, leading to early proposals that the identified complex repressed gene expression through a cryptic ribonuclease activity.  This lead to the particle initially being named the ‘prosome’ \citep{kremp86, schmid84}.  Subsequent analyses of these preparations by SDS-PAGE and electron microscopy revealed the signature ladder of $\alpha$- and $\beta$-subunits at 20-35 kDa, as well as their barrel-like architecture (\ref{fig:emstructure} A and B; \citep{baumeister88, kremp86, schmid84}).

@@

Purification of the 20S fraction from HeLa cells followed by SDS-PAGE also gave rise to this stereotypical protein banding pattern and shape \citep{schmid84}, and this was followed shortly thereafter by the first description of plant prosomes, purified from tobacco leaf extracts using similar sedimentation protocols in ATP-free buffers \citep{kremp86}.  In these later cases, the purified preparations had strong peptidase activity but little to no RNase activity, thus leading to the conclusion that the CP is actually a protease.  Once its true function in protein turnover was confirmed, the moniker for the particle was changed to ‘proteasome’ \citep{arrigo88}.

@@

The complete 26S proteasome (i.e. the CP capped at one or both ends by the RP) was subsequently discovered by the purification of ubiquitin conjugate-degrading activity from rabbit reticulocytes \citep{hough86}.  While it had been well established that major catabolic processes in animal cells involved the ATP-dependent proteolysis of selective substrates \citep{etlinger77}, the enzyme(s) responsible for this activity had yet to be identified.  Taking advantage of the new ability to synthesize ubiquitylated substrates such as 125I-labelled ubiquitin-lysozyme conjugates \citep{hough86-1xVPf}, a protocol was developed to purify the responsible ATP-dependent protease.  Through a series of anion exchange and size exclusion chromatography steps followed by glycerol gradient sedimentation, all of which were performed in ATP-containing buffers, the responsible activity was isolated \citep{hough86, hough87}. The active enzyme turned out to be the 20S proteasome (i.e. the CP) along with a number of additional polypeptides, which together formed a 26S particle, thus providing the first direct link between ubiquitylation and a protease \citep{ganoth88, hough87, waxman87}.  SDS-PAGE analysis of these preparations identified a host of new polypeptides in the 35-100 kDa range in addition to the known CP subunits, which were later shown to comprise a second stable complex, the RP.  Shortly thereafter, the RP was demonstrated to have ATPase activities attributable to the RPT subunits, which help in substrate unfolding and maintaining CP-RP association \citep{armon90}.  Electron microscopic images of the full 26S particle then revealed its diagnostic quaternary structure in which the CP is capped by one or two RPs that sit over the axial pores for substrate entry (\ref{fig:emstructure} C; \citep{peters91, yoshimura93}).

The existence of a similar 26S proteasome in plants was initially implied by the detection of an ATP-dependent activity in oat and wheat germ extracts capable of degrading ubiquitylated proteins \citep{hatfield89, vierstra88}.  This was followed some years later by the first isolation of a complete plant 26S proteasome holocomplex from spinach leaves \citep{fujinami94}.  As with the mammalian forms, purification was achieved by anion exchange and size exclusion chromatography, followed by glycerol gradient centrifugation, all in the presence of ATP to stabilize the CP-RP association.  These spinach preparations were, like their rabbit reticulocyte counterparts, able to rapidly degrade ubiquitylated substrates in an ATP-dependent manner, and further analysis by native-PAGE, SDS-PAGE and electron microscopy revealed the complete subunit composition and “caterpillar-like” structure of the plant particle (\ref{fig:emstructure} D; \citep{fujinami94}).  Similar purifications were successful using rice suspension culture cells and garlic cloves \citep{malik04, yanagawa99}, which were accompanied by the first demonstrations that proteasome inhibitors designed for their mammalian counterparts were effective with the plant particles, suggesting very similar enzymatic mechanisms \citep{ozaki92, woffenden98}.

Despite its prevalence as a genetic model, purification of the 26S proteasome from the flowering plant \textit{Arabidopsis thaliana} was not reported until several years after other plant species \citep{yang04}.  First protocols involved differential PEG precipitation followed by anion exchange and size exclusion chromatography, with the latter exploiting the large size of the holoprotease.  More recently, an improved one-step affinity method was developed \citep{book10}, based on the strategies that had been successfully employed in yeast \citep{leggett05}.  Here, epitope-tagged proteasomes were generated by genetically replacing the subunit PAG1 with a variant bearing a C-terminal FLAG tag; this tagged particle could then be purified with appropriate affinity matrices, and released in non-denaturing conditions with FLAG peptide.  This approach enables rapid and robust purification of the whole 26S proteasome complex when performed in the presence of ATP, or enables purification of the CP sub-particle when performed in the absence of ATP \citep{book10}. This affinity method has considerable advantages compared to previous, conventional chromatographic approaches \citep{yang04} as it is both faster, more reliable and produces higher yields per gram of tissue (\textasciitilde{}6 µg/g). This milder more rapid technique also prevents breakdown of some subunits, in particular RPN10, which is sensitive to post-homogenization proteolysis \citep{yang04}.  One caveat is that the epitope tag, given its exposed position and flexible structure, might be sensitive to proteolytic cleavage following tissue homogenization.  For the PAG1-FLAG protocol, chymostatin was found to effectively block the interfering protease \citep{book10}. An example of such preparations analyzed by SDS-PAGE followed by immunoblotting with antibodies against several proteasome subunits, is shown in \ref{fig:pag1flagap}.

@@

\subsection{20S Core Protease Assembly}
	 In yeast and mammals five of the seven $\beta$-subunits ($\beta$1, $\beta$2, $\beta$5, $\beta$6, and $\beta$7) are synthesized as an immature propeptide form with an N-terminal extension. These $\beta$-subunits undergo autolytic cleavage exposing a catalytically active N-terminal threonine for the $\beta$1, $\beta$2, $\beta$5 subunits upon CP maturation.  \citep{gu14, lee90}. This propeptide prevents inactivation of the mature form’s N-terminal threonine by N-terminal acetylation, and also prevents premature activation of peptidase activities \citep{arendt99, arendt97}. While this propeptide processing occurs in \textit{Arabidopsis}, and likely other plants, its role in plant proteasome assembly remains to be investigated \citep{book10}. In yeast, complexes containing these immature propeptide forms were first identified in 13-16S fractions (smaller than the 20S fraction) suggesting that distinct assembly steps may occur \citep{frentzel94}. Indeed in yeast these propeptides were found to be important in the assembly of the 20S CP as deletion of the $\beta$5 propeptide was shown to be lethal by preventing assembly of the CP, and deletion of the $\beta$2 propeptide accumulated aberrant complexes \citep{jager99}. Several of these propeptides were found to interact with the first dedicated assembly chaperone discovered, Ump1, which was identified in a screen for null mutants defective in ubiquitin-mediated proteolysis in yeast \citep{li07, ramos98}. Ump1 along with other dedicated assembly chaperones aide in the assembly of 15S complexes, or proteasome half barrels containing $\alpha$-subunits 1-7, and $\beta$-subunits 1-6, noticeably lacking the $\beta$7 subunit \citep{li07, marques07}. Similar investigations of the mammalian $\beta$5 propeptide enabled identification of 15S half-barrels that also contained Ump1 and a similar complement of $\alpha$- and $\beta$-subunits also lacking the $\beta$7 subunit \citep{witt00}. Later studies would go on to show that in both mammals and yeast, $\beta$7 is the last subunit to enter the $\beta$-ring, resulting in Ump1 being degraded as the first substrate of nascent CP \citep{hirano08, ramos98}. Intriguingly, the alternative capping particle, PA200 co-purified with these 15S half-barrels suggesting that PA200 may be involved in proteasome assembly (see CP assembly summary in \ref{fig:cpassembly}) \citep{li07, marques07}.

@@

	An orthologous system to the yeast Pba1-4 system was elucidated in mammals. PAC1 and PAC2 were first identified in affinity preparations of an immature propeptide form of the $\beta$1 subunit, and like their yeast counterparts (Pba1 and 2) were found to form a stable heterodimer \citep{hirano05}. Consistent with their role in assembly they were found to co-fractionate with immature $\beta$-subunits, and Ump1 \citep{hirano05}. Overexpression of these PAC1 and PAC2 chaperones yielded increases in half- barrel precursors, while knockdown by siRNA impaired assembly of the CP complex \citep{hirano05}. An additional chaperone involved in assembling the $\alpha$-ring was identified in PAC1 immunopurifications in human cell lines, and was termed PAC3 \citep{hirano06}. siRNA knockdown of PAC3 attenuated $\alpha$-ring formation and gave rise to the accumulation of free $\alpha$-subunits, suggesting that PAC3 played a role in stabilization of the CP $\alpha$-ring \citep{hirano06}. Mammalian PAC4 was identified as an ortholog to yeast Pba4 through PSI-BLAST analyses \citep{kusmierczyk08} and together with PAC4 forms a heterodimer, like it’s yeast Pba3/4 counterpart \citep{le07}.

	The role of the PAC1/2 and PAC3/4 heterodimers in assembly, along with the exact ordered incorporation of $\beta$-subunits was aided by siRNA knockdown of specific $\beta$-subunits, and studying the resulting assembly products \citep{hirano08}. Through this analysis Hirano \textit{et al}. clearly delineated PAC3/4’s role in early assembly, and found that PAC3 and PAC4 must leave the assembling half-barrel before the addition of the $\beta$3 subunit. While this was demonstrated in mammals, observations of the yeast Pba3/4 heterodimer suggest a similar model in which the Pba3/4 heterodimer also remains bound to the $\alpha$-ring until the $\beta$2 subunit binds \citep{kusmierczyk08, yashiroda08}. While Pba(PAC)3/4 act early in assembly, Pba(PAC)1/2 have the propensity to bind the mature proteasome, and stay bound even after the incorporation of $\beta$-subunits.

>>>>

Adapted in part from: R.S. Marshall*, D.C. Gemperline*, and R.D. Vierstra (2015). Purification of 26S proteasomes and their sub-complexes from plants. \textit{Methods Mol. Biol.} DOI: 10.1007/978-1-4939-6533-5 \\

\noindent
*These authors contributed equally to the \textit{Methods Mol. Biol.} work. \\

\noindent
This chapter has expanded sections on the proteasomes role in plant biology, proteasome assembly, and proteasome isotypes. \\

\section{The Ubiquitin 26S Proteasome System (UPS)}
	Selective proteolysis in plants plays a critical role in both regulating growth and development, and maintaining cellular homeostasis \citep{nelson14, smalle04, vierstra93, vierstra09}.  One of the principle pathways for protein degradation in plants and other eukaryotes is the ubiquitin-26S proteasome system (UPS), which involves the attachment of polyubiquitin chains to target proteins followed by their recognition and degradation by the 26S proteasome, an exquisitely designed proteolytic machine \citep{bhattacharyya14, finley09, vierstra09}.  The UPS is highly conserved across all eukaryotes; it was first elucidated by elegant work in rabbit reticulocyte lysates \citep{ciechanover80, ciechanover80-frAQB, etlinger77, hershko80, wilkinson80}, and was subsequently identified in other animals, yeast and plants \citep{ciechanover84, finley84, finley87, glotzer91, hochstrasser91, shanklin87}.  Ubiquitin conjugation to target proteins is accomplished through a highly polymorphic, ATP-dependent cascade involving the sequential action of three enzyme classes, termed the E1 ubiquitin-activating enzymes, E2 ubiquitin-conjugating enzymes, and E3 ubiquitin-protein ligases \citep{berndsen14, smalle04, vierstra09} (Figure \ref{fig:upscycle}).  Selectivity in ubiquitylation is typically driven by the E3 family, which has dramatically expanded during plant evolution to include well over a thousand variants in Arabidopsis and other plant species \citep{hua13, hua11}. Indeed 6\% of the Arabidopsis proteome is devoted to UPS \citep{smalle04}.

@@

	Additionally, the UPS also aides in the regulation of key plant hormonal responses including auxin, abscisic acid, jasmonic acid (JA), giberellins (GA), and ethylene. (reviewed in \citep{santner10}). Key hormone response factors are degraded by a variety of E3 complexes including TIR1 involved in auxin regulation \citep{dharmasiri05}, COI1, involved JA sensing \citep{katsir08}, and GID1 involved in GA responses \citep{murase08}. An elegant mechanism requires the direct binding of these key hormones either to their respective E3 complexes, or key proteins associated with these E3's to promote the efficient ubiquitylation and subsequent degradation of their hormone response factors \citep{shabek14}.

@@

\section{The 26S Proteasome}
	The 26S proteasome is the central enzymatic effector of the UPS, and is a 2.5 MDa particle located in the cytosol and nucleus of eukaryotic cells.  It is composed of two functionally distinct sub-complexes; the 20S core protease (CP) that houses the proteolytic active sites, and the 19S regulatory particle (RP) that recognizes appropriate substrates (Figure \ref{fig:proteasome} A and B; \citep{bhattacharyya14, finley09, lander12, lasker12, unverdorben14}).

@@

The CP has a barrel shape generated by four stacked hetero-heptameric rings, which contain seven $\alpha$-subunits or seven $\beta$-subunits (termed PAA-PAG and PBA-PBG, respectively, in \textit{Arabidopsis}) in an $\alpha$1-7/$\beta$1-7/$\beta$1-7/$\alpha$1-7 configuration (Figure \ref{fig:cpdetails} A (side view- stacked rings), C (top view- $\alpha$-subunits) and D (top view- $\beta$-subunits).

@@

Upon assembly, a central chamber is formed at the $\beta$-ring interface that houses six peptidase catalytic sites provided by the $\beta$1 (PBA), $\beta$2 (PBB), and $\beta$5 (PBE) subunits \citep{arendt97, heinemeyer97} (Figure \ref{fig:cpdetails} B and E).  The active sites involve a catalytic triad, one residue of which is an N-terminal threonine that becomes exposed during CP assembly.  Collectively these peptidases can cleave a broad range of protein sequences with peptidylglutamyl-peptide hydrolyzing (PGPH) ($\beta$1), trypsin-like ($\beta$2), and chymotryptic-like ($\beta$5) activities \citep{arendt97, groll99} (Figure \ref{fig:cpdetails} E).  The $\alpha$-rings create two antechambers with narrow opposing axial pores that are gated by extensions at the N-terminus of several subunits (Figure \ref{fig:cpdetails} B) \citep{groll00, ruschak10}.  Through this distinctive architecture, the CP acts as a self-compartmentalized protease that will only degrade polypeptides that are deliberately recognized, unfolded, and imported into the $\beta$-ring chamber.

	The CP is capped at one or both ends by the RP, which sits on top of the axial pores.  The RP provides activities for recognition of ubiquitylated proteins, substrate unfolding and import, and release of the ubiquitin moieties before substrate degradation (Figure \ref{fig:proteasome}).  Its binding to the CP is stabilized by ATP, which is thus a necessary ingredient for purifying intact 26S proteasomes \citep{smith05}. The RP itself consists of two sub-complexes; the base, which contains a hexameric ring of AAA-ATPases (RPT1-6) plus two non-ATPase subunits, RPN1 and RPN2; and the lid, which is composed of an additional 11 non-ATPase subunits, RPN3, RPN5-13 and DSS1/SEM1 (Figure \ref{fig:proteasome}; \citep{bhattacharyya14, book10, finley09, glickman98-c8Wsa, russell13}.  This lid/base demarcation was first revealed by the absence of lid subunits in proteasomes isolated from a \textit{Δrpn10} yeast deletion strain, and it was hence thought that RPN10 helps enforce binding of the lid to the base \citep{glickman98}.  However, more recent structural studies have demonstrated that RPN10 has a more indirect stabilizing role via its interaction with RPN9 \citep{lander12}.  The ring of RPT subunits in the base promotes substrate unfolding through ATP hydrolysis, and gates the $\alpha$-ring axial pores through repositioning of the CP $\alpha$-subunit extensions \citep{köhler01, rabl08, smith05}.  The N-terminal regions from proximal RPT pairs intertwine to create three spokes onto which most RPN subunits are scaffolded (Figure \ref{fig:rpdetails} A; \citep{beck12}).

@@

Several RPT subunits harbor C-terminal \textit{Hydrophobic-Tyrosine-Any} amino acid (HbYX) motifs important for their binding to the $\alpha$-subunits in the CP. These C-terminal HbYX and HbYX-like extensions typically intercalate between adjacent $\alpha$-subunits (Figure \ref{fig:rpdetails} B). The RPN6 subunit acts as a molecular clamp to tether the RP onto the CP (Figure \ref{fig:rpdetails} A) \citep{pathare12}.

\section{Proteasome Purification Strategies}
	Even before the realization that the 26S proteasome is a protease, sub-particles of the complexes were described.  The first reports of proteasomes used avian erythroblast preparations enriched by differential ultracentrifugation followed by fractionation through a sucrose gradient \citep{schmid84}.  These 20S fractions isolated in the absence of added ATP were found to inhibit mRNA translation in a cell-free system, leading to early proposals that the identified complex repressed gene expression through a cryptic ribonuclease activity.  This lead to the particle initially being named the ``prosome'' \citep{kremp86, schmid84}.  Subsequent analyses of these preparations by SDS-PAGE and electron microscopy revealed the signature ladder of $\alpha$- and $\beta$-subunits at 20-35 kDa, as well as their barrel-like architecture (Figure \ref{fig:emstructure} A and B; \citep{baumeister88, kremp86, schmid84}).

@@

Purification of the 20S fraction from HeLa cells followed by SDS-PAGE also gave rise to this stereotypical protein banding pattern and shape \citep{schmid84}, and this was followed shortly thereafter by the first description of plant prosomes, purified from tobacco leaf extracts using similar sedimentation protocols in ATP-free buffers \citep{kremp86}.  In these later cases, the purified preparations had strong peptidase activity but little to no RNase activity, thus leading to the conclusion that the CP is actually a protease.  Once its true function in protein turnover was confirmed, the moniker for the particle was changed to ``proteasome'' \citep{arrigo88}.

@@

The complete 26S proteasome (i.e. the CP capped at one or both ends by the RP) was subsequently discovered by the purification of ubiquitin conjugate-degrading activity from rabbit reticulocytes \citep{hough86}.  While it had been well established that major catabolic processes in animal cells involved the ATP-dependent proteolysis of selective substrates \citep{etlinger77}, the enzyme(s) responsible for this activity had yet to be identified.  Taking advantage of the new ability to synthesize ubiquitylated substrates such as 125I-labelled ubiquitin-lysozyme conjugates \citep{hough86-1xVPf}, a protocol was developed to purify the responsible ATP-dependent protease.  Through a series of anion exchange and size exclusion chromatography steps followed by glycerol gradient sedimentation, all of which were performed in ATP-containing buffers, the responsible activity was isolated \citep{hough86, hough87}. The active enzyme turned out to be the 20S proteasome (i.e. the CP) along with a number of additional polypeptides, which together formed a 26S particle, thus providing the first direct link between ubiquitylation and a protease \citep{ganoth88, hough87, waxman87}.  SDS-PAGE analysis of these preparations identified a host of new polypeptides in the 35-100 kDa range in addition to the known CP subunits, which were later shown to comprise a second stable complex, the RP.  Shortly thereafter, the RP was demonstrated to have ATPase activities attributable to the RPT subunits, which help in substrate unfolding and maintaining CP-RP association \citep{armon90}.  Electron microscopic images of the full 26S particle then revealed its diagnostic quaternary structure in which the CP is capped by one or two RPs that sit over the axial pores for substrate entry (Figure \ref{fig:emstructure} C; \citep{peters91, yoshimura93}).

The existence of a similar 26S proteasome in plants was initially implied by the detection of an ATP-dependent activity in oat and wheat germ extracts capable of degrading ubiquitylated proteins \citep{hatfield89, vierstra88}.  This was followed some years later by the first isolation of a complete plant 26S proteasome holocomplex from spinach leaves \citep{fujinami94}.  As with the mammalian forms, purification was achieved by anion exchange and size exclusion chromatography, followed by glycerol gradient centrifugation, all in the presence of ATP to stabilize the CP-RP association.  These spinach preparations were, like their rabbit reticulocyte counterparts, able to rapidly degrade ubiquitylated substrates in an ATP-dependent manner, and further analysis by native-PAGE, SDS-PAGE and electron microscopy revealed the complete subunit composition and “caterpillar-like” structure of the plant particle (Figure \ref{fig:emstructure} D; \citep{fujinami94}).  Similar purifications were successful using rice suspension culture cells and garlic cloves \citep{malik04, yanagawa99}, which were accompanied by the first demonstrations that proteasome inhibitors designed for their mammalian counterparts were effective with the plant particles, suggesting very similar enzymatic mechanisms \citep{ozaki92, woffenden98}.

Despite its prevalence as a genetic model, purification of the 26S proteasome from the flowering plant \textit{Arabidopsis thaliana} was not reported until several years after other plant species \citep{yang04}.  First protocols involved differential PEG precipitation followed by anion exchange and size exclusion chromatography, with the latter exploiting the large size of the holoprotease.  More recently, an improved one-step affinity method was developed \citep{book10}, based on the strategies that had been successfully employed in yeast \citep{leggett05}.  Here, epitope-tagged proteasomes were generated by genetically replacing the subunit PAG1 with a variant bearing a C-terminal FLAG tag; this tagged particle could then be purified with appropriate affinity matrices, and released in non-denaturing conditions with FLAG peptide.  This approach enables rapid and robust purification of the whole 26S proteasome complex when performed in the presence of ATP, or enables purification of the CP sub-particle when performed in the absence of ATP \citep{book10}. This affinity method has considerable advantages compared to previous, conventional chromatographic approaches \citep{yang04} as it is both faster, more reliable and produces higher yields per gram of tissue (\textasciitilde{}6 µg/g). This milder more rapid technique also prevents breakdown of some subunits, in particular RPN10, which is sensitive to post-homogenization proteolysis \citep{yang04}.  One caveat is that the epitope tag, given its exposed position and flexible structure, might be sensitive to proteolytic cleavage following tissue homogenization.  For the PAG1-FLAG protocol, chymostatin was found to effectively block the interfering protease \citep{book10}. An example of such preparations analyzed by SDS-PAGE followed by immunoblotting with antibodies against several proteasome subunits, is shown in Figure \ref{fig:pag1flagap}.

@@

\subsection{20S Core Protease Assembly}
	 In yeast and mammals five of the seven $\beta$-subunits ($\beta$1, $\beta$2, $\beta$5, $\beta$6, and $\beta$7) are synthesized as an immature propeptide form with an N-terminal extension. These $\beta$-subunits undergo autolytic cleavage exposing a catalytically active N-terminal threonine for the $\beta$1, $\beta$2, $\beta$5 subunits upon CP maturation.  \citep{gu14, lee90}. This propeptide prevents inactivation of the mature form's N-terminal threonine by N-terminal acetylation, and also prevents premature activation of peptidase activities \citep{arendt99, arendt97}. While this propeptide processing occurs in \textit{Arabidopsis}, and likely other plants, its role in plant proteasome assembly remains to be investigated \citep{book10}. In yeast, complexes containing these immature propeptide forms were first identified in 13-16S fractions (smaller than the 20S fraction) suggesting that distinct assembly steps may occur \citep{frentzel94}. Indeed in yeast these propeptides were found to be important in the assembly of the 20S CP as deletion of the $\beta$5 propeptide was shown to be lethal by preventing assembly of the CP, and deletion of the $\beta$2 propeptide accumulated aberrant complexes \citep{jager99}. Several of these propeptides were found to interact with the first dedicated assembly chaperone discovered, Ump1, which was identified in a screen for null mutants defective in ubiquitin-mediated proteolysis in yeast \citep{li07, ramos98}. Ump1 along with other dedicated assembly chaperones aide in the assembly of 15S complexes, or proteasome half barrels containing $\alpha$-subunits 1-7, and $\beta$-subunits 1-6, noticeably lacking the $\beta$7 subunit \citep{li07, marques07}. Similar investigations of the mammalian $\beta$5 propeptide enabled identification of 15S half-barrels that also contained Ump1 and a similar complement of $\alpha$- and $\beta$-subunits also lacking the $\beta$7 subunit \citep{witt00}. Later studies would go on to show that in both mammals and yeast, $\beta$7 is the last subunit to enter the $\beta$-ring, resulting in Ump1 being degraded as the first substrate of nascent CP \citep{hirano08, ramos98}. Intriguingly, the alternative capping particle, PA200 co-purified with these 15S half-barrels suggesting that PA200 may be involved in proteasome assembly (see CP assembly summary in Figure \ref{fig:cpassembly}) \citep{li07, marques07}.

@@

	An orthologous system to the yeast Pba1-4 system was elucidated in mammals. PAC1 and PAC2 were first identified in affinity preparations of an immature propeptide form of the $\beta$1 subunit, and like their yeast counterparts (Pba1 and 2) were found to form a stable heterodimer \citep{hirano05}. Consistent with their role in assembly they were found to co-fractionate with immature $\beta$-subunits, and Ump1 \citep{hirano05}. Overexpression of these PAC1 and PAC2 chaperones yielded increases in half- barrel precursors, while knockdown by siRNA impaired assembly of the CP complex \citep{hirano05}. An additional chaperone involved in assembling the $\alpha$-ring was identified in PAC1 immunopurifications in human cell lines, and was termed PAC3 \citep{hirano06}. siRNA knockdown of PAC3 attenuated $\alpha$-ring formation and gave rise to the accumulation of free $\alpha$-subunits, suggesting that PAC3 played a role in stabilization of the CP $\alpha$-ring \citep{hirano06}. Mammalian PAC4 was identified as an ortholog to yeast Pba4 through PSI-BLAST analyses \citep{kusmierczyk08} and together with PAC4 forms a heterodimer, like its yeast Pba3/4 counterpart \citep{le07}.

	The role of the PAC1/2 and PAC3/4 heterodimers in assembly, along with the exact ordered incorporation of $\beta$-subunits was aided by siRNA knockdown of specific $\beta$-subunits, and studying the resulting assembly products \citep{hirano08}. Through this analysis Hirano \textit{et al}. clearly delineated PAC3/4's role in early assembly, and found that PAC3 and PAC4 must leave the assembling half-barrel before the addition of the $\beta$3 subunit. While this was demonstrated in mammals, observations of the yeast Pba3/4 heterodimer suggest a similar model in which the Pba3/4 heterodimer also remains bound to the $\alpha$-ring until the $\beta$2 subunit binds \citep{kusmierczyk08, yashiroda08}. While Pba(PAC)3/4 act early in assembly, Pba(PAC)1/2 have the propensity to bind the mature proteasome, and stay bound even after the incorporation of $\beta$-subunits.

	A model of yeast and mammalian assembly adapted from Hirano \textit{et al}. is as follows (Figure \ref{fig:cpassembly}).  After the $\alpha$-ring is formed, with the help of Pba(PAC)1/2 and Pba(PAC)3/4, the $\beta$2 subunit is incorporated along with its C-terminal tail that helps to recruit the $\beta$3 subunit \citep{hirano08}. Pba(PAC)3/4 is released upon $\beta$3 subunit binding \citep{hirano08}. Ump1 stabilizes the incorporation of additional $\beta$-subunits by binding their inherently disordered propeptide regions \citep{li07}. $\beta$4, $\beta$5, $\beta$6, and $\beta$1 are then incorporated \citep{hirano08}. At this stage PA200 may help stabilize this precursor; however, its exact role in assembly is poorly understood.  Two 15S half proteasomes come together, and finally the $\beta$7 subunit is incorporated \citep{hirano08}. The $\beta$-subunit propeptides are then cleaved exposing the catalytically active N-terminal threonines and Ump1 is degraded becoming the first substrate of the nascent CP \citep{li07}. Pba(PAC)1/2 are then released, or can bind mature proteasomes until activation by the RP \citep{hirano08}. Taken together these data are consistent with a model in which the heterodimeric pair Pba(PAC)3/4 acts at early stages to assemble free $\alpha$-subunits into an $\alpha$-ring while preventing the formation of aberrant $\alpha$-subunit complexes \citep{hirano08}. In contrast, the heterodimeric pair Pba(PAC)1/2 acts at both early and later stages in proteasome assembly and has the ability to bind mature CP \citep{wani16}.
